Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-17 ICML-submission/introduction.tex
Fix introdcution

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Classification with Bandit Feedback \chicheng{Shall we emphasize linear here?}}
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\FOR{$t=1,2,\dots,T$}
\STATE Adversary chooses example $(x_t, y_t)$, where $x_t$ is revealed to the learner.
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$.}
\STATE{Observe feedback $z_t \in \{0,1\}$, \\ \qquad where $z_t = \begin{cases}
0 & \text{if $\widehat y_t = y_t$,} \\
1 & \text{if $\widehat y_t \neq y_t$.}
\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{protocol}

@@

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}, formally
stated in
Definitions~\ref{definition:weak-linear-separability}~and~\ref{definition:strong-linear-separability}.
The standard notion of multiclass linear separability~\citep[See e.g.][]{Crammer-Singer-2003}
corresponds to the weak
linear separability. For binary classification, the weak and strong separability
are identical. However, they differ for multiclass classification with $K \ge 3$
classes. For multiclass classification with $K$ classes, weak linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane.

In the full-information feedback setting, it is well known that if all examples
have norm at most $R$ and are weakly linearly separable with a margin $\gamma$,
then the \textsc{Multiclass Perceptron} algorithm makes at most $\lfloor
2(R/\gamma)^2 \rfloor$ mistakes.
\chicheng{I just noted that Theorem 3 of~\cite{Crammer-Singer-2003} actually
has exactly the same upper bound - but it might still make sense to include the
proof here.}
The guarantees provided by \textsc{Multiclass
Perceptron} is very satisfying, since it achieves an information-theoretically
optimal mistake upper bound, while having low time and memory
complexities.~\footnote{These results are standard in the literature. For
completeness, we present these results along with their proofs in
Appendix~\ref{section:multiclass-perceptron-proofs} in the supplementary
material.}

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their proof of
Theorem 2 is incorrect: their Lemma 4 cannot be directly applied to the bandit
setting, as the algorithm only makes updates on a subset of the $w_i$'s.} We
design a simple and efficient algorithm, namely
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
that has an expected number of mistakes at most $O(K (R/\gamma)^2))$.
The memory complexity and per-round time complexity are both $O(dK)$.
The algorithm can be viewed as running $K$ copies of the \textsc{Binary
Perceptron} algorithm, one copy for each class. Intuitively speaking, the $O(K)$
multiplicative factor in the mistake bound is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We prove
that any (possibly randomized) algorithm must make $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.
%$(K-1) \lfloor 4(R/\gamma)^2\rfloor$

>>>>

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Linear Classification with Bandit Feedback}
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\FOR{$t=1,2,\dots,T$}
\STATE Adversary chooses example $(x_t, y_t)$, where $x_t$ is revealed to the learner.
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$.}
\STATE{Observe feedback $z_t \in \{0,1\}$, \\ \qquad where $z_t = \begin{cases}
0 & \text{if $\widehat y_t = y_t$,} \\
1 & \text{if $\widehat y_t \neq y_t$.}
\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{protocol}

@@

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}, formally
stated in
Definitions~\ref{definition:weak-linear-separability}~and~\ref{definition:strong-linear-separability}.
The standard notion of multiclass linear
separability~\citep{Crammer-Singer-2003} corresponds to the weak linear
separability. For binary classification, the weak and strong separability are
identical. However, they differ for multiclass classification with $K \ge 3$
classes. For multiclass classification with $K$ classes, weak linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane.

In the full-information feedback setting, it is well known
\citep{Crammer-Singer-2003} that if all examples have norm at most $R$ and are
weakly linearly separable with a margin $\gamma$, then the \textsc{Multiclass
Perceptron} algorithm makes at most $\lfloor 2(R/\gamma)^2 \rfloor$ mistakes.
The guarantees provided by \textsc{Multiclass Perceptron} is very satisfying,
since it achieves an information-theoretically optimal mistake upper bound,
while having low time and memory complexities.~\footnote{These are
folklore results. For completeness, we present these results along
with their proofs in Appendix~\ref{section:multiclass-perceptron-proofs} in the
supplementary material.}

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their proof of
Theorem 2 is incorrect: their Lemma 4 cannot be directly applied to the bandit
setting, as the algorithm only makes updates on a subset of the $w_i$'s.} We
design a simple and efficient algorithm, namely
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
that has an expected number of mistakes at most $O(K (R/\gamma)^2))$.
The memory complexity and per-round time complexity are both $O(dK)$.
The algorithm can be viewed as running $K$ copies of the \textsc{Binary
Perceptron} algorithm, one copy for each class. Intuitively speaking, the $O(K)$
multiplicative factor in the mistake bound is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We prove
that any (possibly randomized) algorithm must make $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.
